### 2013-09-12 paper_draft.tex
minor changes to philipps revision

%\subsection*{Survey issues}
%
%As the survey was taken online by voluntary participants and was mainly spread in the personal genetics community, the results do not reflect the general population, but over-represent those people most likely to be interested in a project such as openSNP: customers of DTC genetic testing companies and people with a high interest in biology. 
%

@@

%Our survey has shown that 
As people are concerned about their privacy and fear that stakeholders like employers, insurance companies, governments or advertisers might misuse the information \cite{Wolinsky2005}, policy makers start to react to those changes by having introduced laws like the 
\textit{Genetic Information Non-Discrimination Act} (GINA) in the United States or the \emph{Gendiagnostikgesetz} (GenDG) in Germany to minimize the impact of
widely available genetic information. DTC genetic testing companies themselves also try to educate their customers about the risks of releasing genetic data \cite{Lee2009}. Neither GINA nor the GenDG offer a complete protection of genetic discrimination, as certain things, such as life insurances, are not covered by those laws.

@@

\subsection*{GWAS and Open Data}
Although prices of exome or even full genome sequencing are dropping rapidly, GWAS are still considerably cheaper. However, GWAS can only detect correlations of SNPs with those traits and do not allow 
inference on the cause for any correlation. Furthermore, for a statistically sound analysis, GWAS need a large enough sample size. Nevertheless, GWAS are still frequently used and new associations are still being discovered \cite{10.1371.journal.pone.0031470,10.1371.journal.pone.0030309,10.1371.journal.pone.0029848}.

@@

By offering a platform where customers of DTC genetic testing can publish their data openSNP for the general public tries to enable and facilitate the re-use of this already generated data. Allowing interested parties to use the data for studies on their own allows people to perform studies without the need to generate genetic data on their own. Additionally the data can be used to enrich other data sets in order to overcome limited sample sizes, which is especially of interest for rare diseases.

>>>>

%\subsection*{Survey issues}
%
%As the survey was taken online by voluntary participants and was mainly spread in the personal genetics community, the results do not reflect the general population, but over-represent those people most likely to be interested in a project such as openSNP: customers of DTC genetic testing companies and people with a high interest in biology. 
%
Here, we present openSNP, a crowdsourced resource that enables customers of DTC testing companies to share their genotypings with researchers and receive new annotations for their genetic variants. With the number of active users already present on openSNP we have shown that customers of DTC companies are willing to share their data at no cost to researchers around the world and are willing to annotate their data with phenotypes.

@@

%Our survey has shown that 
As people are concerned about their privacy and fear that stakeholders like employers, insurance companies, governments or advertisers might misuse the information \cite{Wolinsky2005}, policy makers start to react to those changes by having introduced laws like the 
\textit{Genetic Information Non-Discrimination Act} (GINA) in the United States or the \emph{Gendiagnostikgesetz} (GenDG) in Germany to minimize the impact of
widely available genetic information. DTC genetic testing companies themselves also try to create online communities - like the 23andMe community forums - that help in educating their customers about the risks of releasing genetic data \cite{Lee2009}. Neither GINA nor the GenDG offer a complete protection of genetic discrimination, as certain things, such as life insurances, are not covered by those laws.

@@

\subsection*{GWAS and Open Data}
Although prices of exome or even full genome sequencing are dropping rapidly, GWAS are still considerably cheaper. However, GWAS can only detect correlations of SNPs with those traits and do not allow 
inference on the cause for any correlation. Furthermore, for a statistically sound analysis, GWAS need a large enough sample size, which often isn't too easy to obtain. Either because generating the needed amount of data still is a cost factor or because it is hard to find enough participants for the case conditions, for example if rare diseases are to be studied. Nevertheless, GWAS are still frequently used and new associations are still being discovered \cite{10.1371.journal.pone.0031470,10.1371.journal.pone.0030309,10.1371.journal.pone.0029848}.

@@

openSNP tries to enable and facilitate the re-use of this already generated data by offering a platform where customers of DTC genetic testing can publish their results into the public domain. Allowing interested parties to use the data for their own research allows scientists to perform studies without the need to generate genetic data sets on their own. Additionally the data can be used to enrich other data sets in order to overcome limited sample sizes, which is especially of interest for rare diseases.
